\documentclass[11pt]{report}

% Loading necessary packages
\usepackage[a4paper,margin=0.75in]{geometry}
\usepackage[english]{babel}
\usepackage{float}
\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\usepackage{pifont}
\usepackage{xcolor}
\usepackage{ulem}
\usepackage{fontawesome}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, calc}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{tcolorbox}
\tcbuselibrary{breakable, skins}
\usepackage{titlesec}
\usepackage{setspace}
\usepackage{microtype}
\usepackage{listings}
\usepackage[switch]{lineno}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{pgfgantt}

% Typography settings
\onehalfspacing
\usepackage{mathptmx}

% Header and Footer Settings
\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\rhead{Certifications To-Do List}
\lhead{Gerald's Escape Plan}
\cfoot{\thepage}

% Define Eisenhower Matrix colors
\definecolor{Do}{rgb}{0.94, 0.6, 0.6}       % Red for urgent and important
\definecolor{Delegate}{rgb}{0.65, 0.95, 0.65} % Green for important but not urgent
\definecolor{Schedule}{rgb}{1.0, 0.85, 0.65}  % Orange for urgent but not important
\definecolor{Eliminate}{rgb}{0.8, 0.8, 0.8}   % Gray for not urgent or important

% Define commands for task categories
\newcommand{\done}{\checkmark}
\newcommand{\pending}{$\square$}
\newcommand{\refine}{$\circlearrowright$}
\newcommand{\issue}{$\triangle$}
\newcommand{\draft}{\faPencil}  
\newcommand{\moved}{\faArrowCircleRight} 

% Define flowchart styles for TikZ
\tikzset{
	startstop/.style = {rectangle, rounded corners, minimum width=2.5cm, minimum height=1cm, text centered, draw=black, fill=red!30, font=\footnotesize},
	process/.style = {rectangle, minimum width=2.5cm, minimum height=1cm, text centered, draw=black, fill=orange!30, font=\footnotesize},
	decision/.style = {rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30, font=\footnotesize},
	arrow/.style = {thick,->,>=stealth}
}

% Custom colors for Eisenhower Matrix
\definecolor{Paid}{rgb}{0.94, 0.6, 0.6}  % Red-ish for urgent and important
\definecolor{Free}{rgb}{0.65, 0.95, 0.65}  % Green for not urgent but important
\definecolor{Trial}{rgb}{1.0, 0.85, 0.65}  % Orange for urgent but not important
\definecolor{Optional}{rgb}{0.8, 0.8, 0.8}  % Gray for not urgent and not important

% Custom commands for Eisenhower Matrix quadrants
\newcommand{\Paid}[1]{\colorbox{Paid}{#1}}          
\newcommand{\Free}[1]{\colorbox{Free}{#1}}   
\newcommand{\Trial}[1]{\colorbox{Trial}{#1}}   
\newcommand{\Optional}[1]{\colorbox{Optional}{#1}}

\begin{document}
	
	\title{\textbf{Certifications To-Do List for Escape Plan}}
	\author{Gerald}
	\date{\today}
	\maketitle
	
\section*{Legend}
\begin{itemize}[label=$\star$]
	\item \makebox[\textwidth][l]{\Free{\textbf{Green}} \dotfill Free}
	\item \makebox[\textwidth][l]{\Trial{\textbf{Yellow}} \dotfill Trial and Trial periods}
	\item \makebox[\textwidth][l]{\Paid{\textbf{Orange}} \dotfill Paid}
\end{itemize}


\section*{Certifications for CV Enhancement}

\subsection*{Essential Certifications}
\begin{itemize}[label=$\Box$, leftmargin=2em]
	\item \Free{\textbf{Bioinformatics Specialization by UC San Diego (Coursera: \$49/month)}}  \\
	Core bioinformatics tools and techniques: genome assembly, sequence alignment.
	
	\item \Paid{\textbf{Nextflow Training (Seqera Labs: \$200)}}  \\
	Workflow orchestration for scalable genomic pipelines.
	
	\item \Trial{\textbf{Microsoft Certified: Azure Data Engineer Associate}}  \\
	Cloud resources for large-scale data processing.
	
	\item \Free{\textbf{Python for Everybody (Coursera, Free with Audit)}}  \\
	Python programming fundamentals for data analysis and automation.
	
	\item \Paid{\textbf{Linux Foundation Certified Engineer (LFCE) (\$300)}}  \\
	Advanced Linux skills for workflow automation.
	
	\item \Trial{\textbf{Google Project Management Certificate}}  \\
	Covers Agile, Scrum, and project lifecycle.
	
	\item \Free{\textbf{Successful Negotiation: Essential Strategies (Coursera, Free)}}  \\
	Communication and persuasion skills.
	
	\item \Paid{\textbf{Business Foundations Professional Certificate (Wharton School, edX: \$149/month)}}  \\
	Essentials of business decision-making, leadership, and financial acumen.
\end{itemize}

\subsection*{Robust Certifications}
\begin{itemize}[label=$\Box$, leftmargin=2em]
	\item \Free{\textbf{Harvard Medical School: Reproducible Science}}  \\
	Emphasizes computational rigor and statistical tools.
	
	\item \Trial{\textbf{DeepLearning.AI TensorFlow Developer Certificate}}  \\
	Master TensorFlow and apply machine learning to bioinformatics datasets.
	
	\item \Paid{\textbf{Google Cloud Professional Data Engineer (\$125 for exam)}}  \\
	Design and deploy scalable data solutions on Google Cloud.
	
	\item \Free{\textbf{EDX: Genomic Data Science by Johns Hopkins}}  \\
	Introduction to tools like Galaxy, Python, and R for genomic analysis.
	
	\item \Paid{\textbf{Certified Kubernetes Administrator (CKA) (\$375)}}  \\
	Containerized workflow management and orchestration for scalable pipelines.
	
	\item \Trial{\textbf{Coursera: Machine Learning by Stanford (Free Trial Available)}}  \\
	Introduction to ML techniques and applications, taught by Andrew Ng.
\end{itemize}

\subsection*{Optional Certifications}
\begin{itemize}[label=$\Box$, leftmargin=2em]
	\item \Paid{\textbf{AWS Certified Solutions Architect – Associate (\$150 for exam)}}  \\
	Cloud-based solution design for scalability and efficiency.
	
	\item \Trial{\textbf{Udemy: Python for Data Science and Machine Learning}}  \\
	Practical Python skills for bioinformatics and data science applications.
	
	\item \Free{\textbf{Intro to Computational Biology (MIT OpenCourseWare)}}  \\
	A foundational course covering algorithms and computational techniques in biology.
	
	\item \Free{\textbf{HubSpot Academy: Inbound Marketing Certification (Free)}}  \\
	Teaches marketing basics, ideal for consulting gigs or promoting personal services.
	
	\item \Paid{\textbf{Introduction to Business Management (University of Edinburgh, edX)}}  \\
	General understanding of business operations and management.
\end{itemize}




	
\end{document}
