\documentclass[11pt,a4paper,sans]{moderncv}

% Adjust margins and layout
\usepackage[a4paper, margin=1in, bottom=1in]{geometry}

% ModernCV themes
\moderncvstyle{banking}
\moderncvcolor{blue}
\nopagenumbers{} % Suppress page numbers

% Required packages for ModernCV
\usepackage{fontspec}
\usepackage{fontawesome5}
\usepackage{academicons}
\usepackage{multicol}

% Fancyhdr for custom footer
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[L]{\normalsize References available only upon request and if shortlisted} 
\setlength{\footskip}{50pt} 
\renewcommand{\headrulewidth}{0pt} 
\renewcommand{\footrulewidth}{0pt}

% Adjust list spacing globally
\usepackage{enumitem}
\setlist[itemize]{topsep=0pt, itemsep=1pt, leftmargin=1.5em}

% Set fonts to Garamond
\setmainfont{EB Garamond}
\setsansfont{EB Garamond}
\setmonofont{Courier New}

% Personal data
\name{Gerald Amiel}{Ballena}
\title{Proofreader for Math, Chemistry, and Biology Tests}
\email{gmballena@up.edu.ph}
\social[github]{GABallena}
\social[linkedin]{gerald-amiel-ballena}
\extrainfo{\aiOrcid \ \href{https://orcid.org/0009-0000-8857-9755}{orcid.org/0009-0000-8857-9755}}

\renewcommand{\makecvtitle}{
	\vspace*{-2em}
	\begin{center}
		{\Huge \textbf{Gerald Amiel Ballena}} \\[0.5em]
		\faEnvelope \ \href{mailto:gmballena@up.edu.ph}{gmballena@up.edu.ph} \quad
		\faGithub \ \href{https://github.com/GABallena}{github.com/GABallena} \quad
		\faLinkedin \ \href{https://linkedin.com/in/gerald-amiel-ballena}{linkedin.com/in/gerald-amiel-ballena} \\[0.5em]
		\faMapMarker \ Metro Manila, Philippines \quad
		\aiOrcid \ \href{https://orcid.org/0009-0000-8857-9755}{orcid.org/0009-0000-8857-9755}
	\end{center}
	\vspace{1.5em}
}

\begin{document}
	
	\makecvtitle
	
\section{Professional Summary}
Detail-oriented and methodical professional with a strong academic background in Bioinformatics, Evolution, and Molecular Biology. Experienced in documenting technical reports, academic papers, and bioinformatics pipelines. Skilled in ensuring clarity, accuracy, and alignment with scientific standards. Strong problem-solving ability with a focus on precision and consistency. Adept at collaborating with multidisciplinary teams to meet high standards of quality in technical and academic writing.

	
	\section{Experience}
	\textbf{2024 – Present} \hfill \textbf{Project Technical Specialist} \\
	\textit{University of the Philippines, College of Public Health} \\[-1em]
	\begin{itemize}
		\item Developed and proofread technical documentation and reports for bioinformatics research.
		\item Ensured that written reports were scientifically accurate and met publication standards.
		\item Reviewed and proofread large datasets and results to ensure accuracy before publication.
	\end{itemize}
	
	\section{Education}
	\subsection{University of the Philippines Diliman}
	\cvitem{Graduated}{July 2022}
	\cvitem{Thesis}{In silico assessment of the association of pathogenicity and metal-resistance potential of \textit{Fusarium} spp.\dotfill \href{https://www.biorxiv.org/content/10.1101/2022.10.12.511937v1}{Pre-print Link}}
	\cvitem{Accomplishments}{DOST ASTHRDP-Scholarship \\ Calculated Effective GPA (MS): 1.72}
	
	\subsection{University of the Philippines Baguio}
	\cvitem{Graduated}{June 2018} 
	\cvitem{Thesis}{\textbf{Bioelectrocatalysis by Novel Electrogenic Alkaliphilic Bacteria \textit{Bacillus} sp. BAB-3442 Using Dual-Chambered Microbial Fuel Cell}  Poster Presentation (\texttt{PSM 47})}
	
	\section{Skills}
	\subsection{Technical Skills}
	\cvitem{Proofreading}{Strong attention to detail with a focus on scientific accuracy in technical writing. Skilled in reviewing and editing academic and technical content, ensuring clarity, and precision.}
	\cvitem{Programming Languages}{Python, R, Bash, Perl \& BioPerl.}
	\cvitem{Bioinformatics Tools}{Kraken2, MetaPhlAn, MEGAHIT, Snakemake, Prokka, Jellyfish, BUSCO, and others.}
	\cvitem{Data Visualization}{ggplot2, matplotlib.}
	\cvitem{High-Performance Computing}{Slurm, Docker.}
	
\subsection{Soft Skills}
\cvitem{Attention to Detail}{Meticulous in documenting technical reports and academic papers, ensuring accuracy.}
\cvitem{Communication}{Strong communication skills to collaborate with teams using appropriate data visualization techniques while remaining flexible enough for feedback from peers for readability.}

	\section{Certifications and Relevant Coursework}
	\subsection{Certifications}
	\cvitem{Fundamentals}{AI Fundamentals \dotfill 2024}
	\cvitem{Fundamentals}{Data Literacy \dotfill 2024}
	
	\subsection{Relevant Courses}
	\cvitem{Specialization}{Biostatistics in Public Health \dotfill 2020}
	\cvitem{Specialization}{Genomic Data Science \dotfill 2024}
	
	\section{Professional Development}
	\cvitem{Ongoing}{Ensemble Methods in Python (Bagging, Boosting, Stacking) \\
		Visualizing Geospatial Data in R \\
		Introduction to AWS}
	
	\section{Scripts and Workflows}
	\cvitem{Key Pipelines and Workflows}{
		Developed bioinformatics pipelines and tools for metagenomics, data analysis, and tool management. Key projects include:
		\begin{itemize}
			\item Automated workflows for sequence analysis, including trimming and taxonomic profiling with Kraken2.
			\item Developed scripts for quality control, data validation, and visualizing bioinformatics data.
		\end{itemize}
	}
	
\end{document}

---

### Key Adjustments:
- **Emphasized proofreading skills**: Highlighted your experience in proofreading academic papers and technical documents, ensuring clarity and accuracy, which is key for this role.
- **Education and Certifications**: Tailored the education and courses section to demonstrate your background in Biology, Chemistry, and Bioinformatics, aligning with the job’s qualifications.
- **Professional Development**: Added relevant ongoing courses and skills development to show your commitment to improving your technical proficiency.

This version of your resume is now tailored for the **Proofreader for Tests** position.
